Less Funding, More Lecturers, and Fewer Professors: \\ \vspace{0.1cm} Stagnating State Funding for Higher Education and its Effect on Faculty

### 2024-05-08 text/sections/01-introduction.tex
Added new robustness check tables.
Newest revision + all code updated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Introduction section
\section{Introduction}
\label{sec:intro}
States fund their public universities with roughly the same amount today as they did in the 1990s, despite enrolment rising by over 50\%.
At the same time, public universities have systematically substituted away from professors towards lecturers, who teach often work on contingent and lower-salaried contracts.
Previous research shows state funding cuts led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether these funding cuts impacted faculty.
In this paper, I show that falling state funding caused a change in the composition of faculty at US public universities, substituting professors for more lecturers.

@@

This paper proceeds as follows.
\autoref{sec:data} describes the data for university funding, data on faculty in Illinois, and recent trends in higher education funding.
\autoref{sec:conceptual} gives the conceptual framework for how state funding may affect faculty.
\autoref{sec:empirics} draws the empirical framework for isolating the causal effects of state funding on faculty, with \autoref{sec:results} presenting the empirical results.
\autoref{sec:discussion} discusses the context and implications for the findings.
\autoref{sec:conclusion} concludes.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Introduction section
%\section{Introduction}
%\label{sec:intro}
States fund their public universities with roughly the same amount today as they did in the 1990s, despite enrolment rising by over 50\%.
At the same time, public universities have systematically substituted away from professors towards lecturers, who teach often work on contingent and lower-salaried contracts.
Previous research shows state funding cuts led to worsening student outcomes \citep{NBERw23736,NBERw27885}, but there is little evidence on whether these funding cuts impacted faculty.
In this paper, I show that falling state funding caused a change in the composition of faculty at US public universities, substituting professors for more lecturers.

@@

This paper proceeds as follows.
\autoref{sec:data} describes the data for university funding, data on faculty in Illinois, and recent trends in higher education funding.
\autoref{sec:conceptual} gives the conceptual framework for how state funding may affect faculty.
\autoref{sec:empirics} draws the empirical framework for isolating the causal effects of state funding on faculty, with \autoref{sec:results} presenting the empirical results.
%\autoref{sec:discussion} discusses the context and implications for the findings.
\autoref{sec:conclusion} concludes.
